Stack Safety as a Security Property

### 2020-06-12 main.tex
Tex for WBCF and treatment of code in confidentiality

\paragraph*{Call Maps and Returns}

A {\it call map} $\callmap$ identifies machine states that represent a
call, abstracting away from the calling convention of any concrete
machine. Formally $\callmap$ is a partial map from machine states to a
natural number, the number of arguments of that call.
%
\[\callmap \in \CALLMAPS = \MACHS \rightharpoonup \mathbb{N}\]

For concreteness, in a typical calling convention, a call map may be
considered to identify the state just before the \(\JAL\), in which
space for arguments is already allocated, but the mechanism is more
general.

A state on which a call map is defined is referred to as a {\it call
  state}, and we consider the call to have returned the first time
control returns to the instruction following the \(\JAL\) with the
stack restored. Formally this is a relation between the states:

@@

A stack safety property is a property of a transition system, a
policy, and a callmap that is decomposed into an integrity property and a
confidentiality property.

@@

\leo{What about instructions? Shouldn't they be LC/HI?}
\sna{We should consider whether applying stack safety to them is right, as
     opposed to a stronger property. Code could change instructions as long as
     they're restored before a callee returns. Shenanigans like with ROP would
     be expected.}
When \(\mpstate{\mach}{\pol}\) is a call state \(\callmap(\mach) = n\)
for some \(n\), we can construct a contour based on the bounds of the
callee's stack frame. Let \(\base \in \WORDS\) be the base of the stack,
and memory below \(\base\) set aside for code.
% Fixed. \leo{Should the first two cases have a $k \not \in \REGS$?}
\[\mathit{Cof}(\mach,n)(\component) =
\begin{cases}
  (\LOWSEC,\HIGHINT) & \text{if } \component \not \in \REGS
                       \text{ and } \component < \base \\
  (\HIGHSEC,\HIGHINT) & \text{if } \component \not \in \REGS
                        \text{ and } \base \leq \component \leq \mach(\SP) - n \\
  (\HIGHSEC,\LOWINT) & \text{if } \component \not \in \REGS
                       \text{ and } \component \geq \mach(\SP) \\
  (\LOWSEC,\LOWINT) & \text{else} \\
\end{cases}\]
The code section below the stack base is low confidentiality and high
integrity, as code should be readable but modifiable.
%
The private state of the callee is the entire area of the memory from stack
base to stack pointer, with the exception of the $n$ locations
corresponding to the arguments of a call. These components are tagged
high confidentiality and high integrity: the callee should be able to
neither read, nor write them.
%
Everything above the stack pointer at the point of the call is
``uninitialized'' memory that can be written into but should not be
read. To protect the callee itself from such reads, these components
are marked high confidentiality and low integrity.
%
Registers and the frame of arguments to the callee are free for the
callee to use, and are therefore tagged with low integrity and low
confidentiality.

%\leo{Shouldn't this be a definition instead of a paragraph?}
%\paragraph{Integrity}

@@

A system satisfies integrity with respect to a callmap \(\callmap\)
if for any initial mpstate \(\mpstatename =
\mpstate{\mach_0}{\mathit{pinit}(\mach_0)}\) and each call state
\(\mpostate{\mach_c}{\pol_c}{\_}\) in the induced trace from \(\mpstatename\)
where \(\callmap(\mach_c) = n\), if we take the prefix from
\(\mpstate{\mach_c}{\pol_c}\) until \(\ret{\mach_c}\) holds on some
\(\mach_p\), \(\mach_p\) matches \(\mach_c\) on each component whose integrity
in \(\mathit{Cof}(\mach_c,n)\) is \(\HIGHINT\).

@@

A system enjoys eager stack confidentiality with respect to a callmap
\(\callmap\) if, for any initial state \(\mpstatename =
\mpstate{\mach_0}{\mathit{pinit}(\mach_0,cm)}\), for each call state
\(\mpostate{\mach_c}{\pol_c}{\_}\) in the induced trace from \(\mpstatename\)
where \(\callmap(\mach_c) = n\) and each variant \(\mach_c'\) such that
\(\mach_c \approx_\contour \mach_c'\) for \(\contour = Cof(\mach_c, n)\),
one of two cases holds.

If the trace from the call reaches a return -- \(\mpstate{\mach_c}{\pol_c}
\hookrightarrow_P \MPT \mpostate{\mach_r}{\pol_r}{\_} | \ret{\mach_c}\) --
then the trace from the variant \(\mach_c'\) reaches a return --
\(\mach_c' \hookrightarrow \machT (\mach_r',\_) | \ret{\mach_c}\).
\(\MPT\) and \(\machT\) are observationally equivalent, \(\pi_o(\MPT) \eqsim
\pi_o(\machT)\). For any component \(\component\) that changed between
\(\mach_c\) and \(\mach_r\), or between \(\mach_c'\) and \(\mach_r'\),
\(\mach_r\) agrees with \(\mach_r'\) on \(\component\).

@@

      \leo{Think more about whether we vary mc or step mc}
      \leo{Also, words}

@@

\sna{Trying to get at why ROP is potentially a problem in non-returning calls
     with an example.}

@@

\paragraph{What about well-bracketed control flow?}

Andrew thinks it's mostly orthogonal...?

@@

\paragraph{Non-stack Resources}

Stack integrity carries an implicit assumption that memory that has
become inaccessible should remain so until the correct return state, and
therefore that a call that never returns fulfills the requirements of
the property. The caller's integrity is not violated until it is able
to execute again and finds that its state has changed.

This assumption might be broken in any number of extended models, such as
the addition of heap memory or global variables. Suppose a function \(f\)
has some private state associated with it, with a static guarantee that
only its instructions access that state. A return oriented programming
attack from a caller could modify the private state using the caller's own
instructions, and even without returning, if the callee calls \(f\) the
modifications will be visible.

A safety property for such an extended model would extend the contour to
determine when the security labels of different classes of components
change, and integrity would mean that when a component is low
confidentiality, it must have the same value as the last time it was low
integrity. Stack integrity is a special case in which a component only
moves from high to low integrity on a return.

@@

      A system enjoys observable integrity with regard to \(\callmap\) if
      for any initial state \(\mach_0\) and each call state \(\mpstatename =
      \mpstate{\mach_c}{\pol_c}\) in the trace induced by
      \(\mpstate{\mach_0}{\mathit{pinit}(\mach_0)}\) where \(\callmap(\mach_c)
      = n\), if there is a return state \(\mpstatename \hookrightarrow \MPT
      \mpostate{\mach_r}{\pol_r}{\_} | \ret{\mach_c}\), the following holds.
      Let \(\contour = \mathit{Cof}(\mach_c,n)\) describe the security levels
      of the call and \(\mach_r' = \mathit{roll}(\contour,\mach_c,\mach_r)\).
      The induced traces \(\mpstate{\mach_r}{\pol_r} \hookrightarrow \MPT\)
      and \(\mach_r' \hookrightarrow \machT'\) must be equivalent up to
      possible policy failure: \(\MPT \sim_P \machT'\).

@@

      A system enjoys observable confidentiality with respect to \(\callmap\)
      if for any initial mpstate \(\mpstatename =
      \mpstate{\mach_0}{\mathit{pinit}(\mach_0)}\), each call state
      \(\mpostate{\mach_c}{\pol_c}{\_}\) in the induced trace from
      \(\mpstatename\) where \(\callmap(\mach_c) = n\), and each variant
      \(\mach_c'\) where \(\mach_c \approx_\contour \mach_c'\) and
      \(\contour = \mathit{Cof}(\mach_c, n)\), one of two cases holds on the
      induced traces from \(\mach_c\) and \(\mach_c'\).

      If \(\mpstate{\mach_c}{\pol_c} \hookrightarrow_P \MPT_1 | \ret{\mach_c}\)
      with \(\MPT_1\) the trace within the callee ending with the machine state
      \(\mach_r\) and policy state \(\pol_r\), the first state to return to the
      caller, then the trace from the variant until return
      \(\mach_c' \hookrightarrow \machT_1 | \ret{\mach_c}\) ends with a return
      state \(\mach_r'\). Let \(\mach_r'' = \mathit{restore}(\contour,\mach_c,
      \mach_c',\mach_r,\mach_r')\), and let the induced traces from the
      returned states be \(\mpstate{\mach_r}{\pol_r} \hookrightarrow_P \MPT_2\)
      and \(\mach_r'' \hookrightarrow \machT_2\). With the callee and caller
      traces joined, the traces are observationally equivalent up to policy
      failures: \(\MPT_1 \cdot \MPT_2 \sim_P \machT_1 \cdot \machT_2\).

@@

>>>>

\paragraph*{Program Structure}

  Our properties and policies are parameterized by several types of annotations
  describing features of a program. A {\em code map} identifies some addresses
  as code and assigns each to a function. A {\em call map} identifies addresses
  that complete a call sequence, abstracting away from any concrete calling
  convention. Likewise a {\em return map} identifies addresses that complete a
  return sequence. An {\em entry map} identifies valid entry points to a
  function.

  Formally a code map is a partial function from addresses to a function
  identifier drawn from the set \(\FUNIDS\).
  %
  \[\codemap \in \CODEMAPS = \WORDS \rightharpoonup \FUNIDS\]

  A call map is a partial function from addresses to a natural number,
  the number of arguments of the call.
  %
  \[\callmap \in \CALLMAPS = \WORDS \rightharpoonup \mathbb{N}\]

  For concreteness, in a typical calling convention, a call map may be
  considered to identify the state just before the \(\JAL\), in which
  space for arguments is already allocated, but the mechanism is more
  general. A state in which a call map is defined on the program counter is
  referred to as a {\it call state}.

  A return map is simply a predicate on addresses and may be considered to
  identify the final jump of the return sequence. An entry map is likewise a
  predicate on addresses.
  %
  \[\retmap \in \RETMAPS \subseteq \WORDS\]
  \[\entmap \in \ENTMAPS \subseteq \WORDS\]

\paragraph*{State Returns}

  We separately identify, for any call state, what is means for a state to
  have correctly returned from a call. We consider the call to have returned
  the first time control returns to the instruction following the \(\JAL\) with
  the stack restored. Formally this is a relation between the states:

@@

  A stack safety property is a property of a transition system, a policy, a
  call map and a code map, that is decomposed into an integrity property and a
  confidentiality property.

@@

  \leo{What about instructions? Shouldn't they be LC/HI?}
  \sna{Now we are dealing with this!}

  When \(\mach_c\) is a call state \(\callmap(\mach(\PCname)) = n\) for some
  \(n\) and \(\mach_c \stepstoobs{\obs} \mach_e\), we can construct a contour
  based on the bounds of the callee's stack frame. Instructions are low
  confidentiality if they belong to the callee and high confidentiality
  otherwise; regardless they are always high integrity.

  The private state of the callee is the entire area of the memory from stack
  base to stack pointer, with the exception of the $n$ locations
  corresponding to the arguments of a call. Private components are marked
  high confidentiality and high integrity: the callee should be able to
  neither read, nor write them.

  Everything above the stack pointer at the point of the call is
  ``uninitialized'' memory that can be written into but should not be
  read. To protect the callee itself from such reads, these components
  are marked high confidentiality and low integrity.

  Registers and the frame of arguments to the callee are free for the
  callee to use, and are therefore tagged with low integrity and low
  confidentiality.

 \[\mathit{Cof}(\mach,\funid,n)(\component) =
  \begin{cases}
    (\LOWSEC,\HIGHINT) & \text{if } \component \in \WORDS
                         \text{ and } \codemap(\component) = \codemap(\mach(\PCname)) \\
    (\HIGHSEC,\HIGHINT) & \text{if } \component \in \WORDS
                          \text{ and } \codemap(\component) = \funid \not = \codemap(\mach(\PCname)) \\
    (\HIGHSEC,\HIGHINT) & \text{if } \component \in \WORDS \text{ and } \codemap(\component) = \bot
                          \text{ and } \component \leq \mach(\SP) - n \\
    (\HIGHSEC,\LOWINT) & \text{if } \component \in \WORDS \text{ and } \codemap(\component) = \bot
                         \text{ and } \component \geq \mach(\SP) \\
    (\LOWSEC,\LOWINT) & \text{else} \\
  \end{cases}\]

@@

  A system satisfies eager stack integrity with respect to a call map
  \(\callmap\) and code map \(\codemap\) if for any initial mpstate
  \(\mpstatename = \mpstate{\mach_0}{\mathit{pinit}(\mach_0)}\) and each
  adjacent pair of a call state \(\mach_c\) and state \(\mach_e\) with policy
  state \(\pol_e\) in the induced trace from \(\mpstatename\), where
  \(\callmap(\mach_c(\PCname)) = n\) and \(\codemap(\mach_e(\PCname)) =
  \funid\), if we take the prefix until a return
  \(\mpstate{\mach_e}{\pol_e} \hookrightarrow_P \MPT | \ret{\mach_c}\) and
  \(\MPT\) ends with some \(\mach_p\), \(\mach_p\) matches \(\mach_c\) on each
  component whose integrity in \(\mathit{Cof}(\mach_c,\funid,n)\) is
  \(\HIGHINT\).

@@

  A system satisfies eager stack confidentiality with respect to a call map
  \(\callmap\) and code map \(\codemap\) if for any initial mpstate
  \(\mpstatename = \mpstate{\mach_0}{\mathit{pinit}(\mach_0)}\) and each
  adjacent pair of a call state \(\mach_c\) and state \(\mach_e\) with policy
  state \(\pol_e\) in the induced trace from \(\mpstatename\), where
  \(\callmap(\mach_c(\PCname)) = n\) and \(\codemap(\mach_e(\PCname)) =
  \funid\), one of two cases holds on the induced traces from
  \(\mpstate{\mach_e}{\pol_e}\) and any variation \(\mach_e'\) with respect
  to the contour \(\mathit{Cof}(\mach_c,\funid,n)\).

  If the trace from the call reaches a return -- \(\mpstate{\mach_e}{\pol_e}
  \hookrightarrow_P \MPT | \ret{\mach_c}\) -- then the trace from the variant
  \(\mach_e'\) reaches a return -- \(\mach_e' \hookrightarrow \machT |
  \ret{\mach_c}\). Let \(\mach_r\) and \(\mach_r'\) be the final states of
  \(\pi_m(\MPT)\) and \(\machT\), respectively. Until their final states,
  \(\MPT\) and \(\machT\) must be observationally equivalent, \(\pi_o(\MPT)
  \cdot \tau \eqsim \pi_o(\machT) \cdot \tau\). For any component
  \(\component\) where \(\mach_e(\component) \not = \mach_r(\component)\) or
  \(\mach_e'(\component) \not = \mach_r'(\component)\),
  \(\mach_r\) agrees with \(\mach_r'\) on \(\component\).

@@

  \leo{Think more about whether we vary mc or step mc}
  \sna{With code, it must be step mc}

@@

@@

  \paragraph{Control Hijacking and Well-bracketed Control Flow}

    Some forms of security rely on protected control flow. In a classic
    example, a system call that launches a missile might be gated behind code
    that checks a password. A piece of malicious code might jump into the middle
    of the sequence and execute the system call without the appropriate password,
    and still obey integrity if the malicious code never returns to its caller.
    Integrity here protects data, not privileges. But such a jump does violate
    eager stack confidentiality if the attacker jumps beyond its own function's
    instructions without performing a designated call or returning to an
    appropriate site.

    A jump to a function's own code does not violate either, even if done using
    nefarious means such as return-oriented programming (ROP). Our model would
    permit such a jump within the function, so privilege is not elevated by
    permitting ROP in this case.

    But, our policies do enforce orthogonal control flow properties that
    protect from calls that enter outside of valid entry points and from all
    invalid returns, even if they would not escalate privilege. Most notably
    {\em well-bracketed control flow}, below.

@@

@@

      A system enjoys observable integrity with regard to \(\callmap\) and 
      \(\codemap\) if for any initial state \(\mach_0\) and each adjacent pair
      of a call state \(\mach_c\) and a state \(\mach_e\) with policy state
      \(\pol_e\) in the trace induced by
      \(\mpstate{\mach_0}{\mathit{pinit}(\mach_0)}\) where
      \(\callmap(\mach_c) = n\) and \(\codemap(\mach_e) = \funid\),
      if \(\mpstate{\mach_e}{\pol_e} \hookrightarrow \MPT | \ret{\mach_c}\)
      with the return state \(\mach_r\) at the end of \(\MPT\), then the
      following holds.

      Let \(\contour = \mathit{Cof}(\mach_c,\funid,n)\) describe the security
      levels of the call and
      \(\mach_r' = \mathit{roll}(\contour,\mach_e,\mach_r)\). The induced
      traces \(\mpstate{\mach_r}{\pol_r} \hookrightarrow \MPT\) and
      \(\mach_r' \hookrightarrow \machT'\) must be equivalent up to
      possible policy failure: \(\MPT \sim_P \machT'\).

@@

      A system enjoys observable confidentiality with respect to \(\callmap\)
      and \(\codemap\) if for any initial state \(\mach_0\), each adjacent pair
      of a call state \(\mach_c\) and a state \(\mach_e\) with policy state
      \(\pol_e\) in the inducted trace from
      \(\mpstate{\mach_0}{\mathit{pinit}(\mach_0)}\), and each variant
      \(\mach_e' \approx_\contour \mach_e'\) where
      \(\contour = \mathit{Cof}(\mach_c, \funid, n)\), one of two cases holds
      on the induced traces from \(\mach_e\) and \(\mach_e'\).

      If \(\mpstate{\mach_e}{\pol_e} \hookrightarrow_P \MPT_1 | \ret{\mach_c}\)
      with \(\MPT_1\) the trace within the callee ending with the machine state
      \(\mach_r\), the first state to return to the caller, and policy state
      \(\pol_r\), then the there is a return state \(\mach_r'\) reached by the
      trace from the variant, \(\mach_c' \hookrightarrow \machT_1 |
      \ret{\mach_c}\). Let \(\mach_r'' = \mathit{restore}(\contour,\mach_e,
      \mach_e',\mach_r,\mach_r')\), and let the induced traces from the
      returned states be \(\mpstate{\mach_r}{\pol_r} \hookrightarrow_P \MPT_2\)
      and \(\mach_r'' \hookrightarrow \machT_2\). With the callee and caller
      traces joined, the traces are observationally equivalent up to policy
      failures: \(\MPT_1 \cdot \MPT_2 \sim_P \machT_1 \cdot \machT_2\).

@@

    \definition{Observable Stack Safety}

      A system enjoys {\em observable stack safety} with regard to a call map
      and a code map if it enjoys both observable integrity and observable
      confidentiality with regard to them.

\section{Well-bracketed Control Flow}

  Our integrity and confidentiality properties concern themselves with data,
  and may hold on a wide range of programs that violate expected control flow.
  But well-bracketed control flow prevents this style of attack by guaranteeing
  that every jump between functions is either a call or a return, and each
  return jumps to the instruction after a corresponding call.

  \definition{Control Separation}

    A system enjoys {\em control separation} with regard to a code map
    \(\codemap\), a call map \(\callmap\), and a return map \(\retmap\)
    if for any initial state \(\mach_0\) and any adjacent pair of states
    \(\mach_1\) and \(\mach_2\) in the induced trace from
    \(\mpstate{\mach_0}{\pinit(\mach_0)}\), if \(\codemap(\mach_1(\PCname))
    \not = \codemap(\mach_2)\), then either \(\callmap(\mach_1(\PCname)) = n\)
    for some \(n\), or \(\retmap(\mach_1)\).

  \definition{Entry Integrity}

    A system enjoys {\em entry integrity} with regard to a call map
    \(\callmap\) and entry map \(\entmap\) if for any initial state
    \(\mach_0\) and any adjacent pair of states \(\mach_c\) and \(\mach_e\)
    in the induced trace from \(\mpstate{\mach_0}{\pinit(\mach_0)}\), if
    \(\callmap(\mach_c(\PCname)) = n\) then \(\entmap(\mach_e)\).

  \definition{Well-bracketed Control Flow}

    Well-bracketed control flow captures the expected behavior of marked
    returns with regard to their call points. The coinductive {\em bracketing}
    predicate, \(\mathit{BR}(\machT)\), is parameterized by a call map
    \(\callmap\) and a return map \(\retmap\):

    \[\frac{\mathit{BR}(\machT) \hspace{.5cm} \callmap(\mach_c(\PCname)) = n
              \hspace{.5cm} \retmap(\mach_c(\PCname))}
           {\mathit{BR}(\mach_c \machT \mach_r)}\]

    \[\frac{\mathit{BR}(\machT) \hspace{.5cm} \callmap(\mach(\PCname)) = \bot}
           {\mathit{BR}(\mach \machT)}\]

    A system enjoys {\em well-bracketed control flow} with regard to a call map
    \(\callmap\) and return map \(\retmap\) if for any initial state
    \(\mach_0\) and any call state and policy state \(\mach_c\) and \(\pol_c\)
    in the induced trace from \(\mpstate{\mach_0}{\pinit(\mach_0)}\), if
    \(\mpstate{\mach_c}{\pol_c} \hookrightarrow_P \MPT \mpostate{\mach'}
    {\pol'}{\obs'} ...\) and \(\MPT\) is bracketed with regard to \(\callmap\)
    and \(\retmap\), then \(\ret{\mach_c,\mach'}\).

  \definition{Standard Stack Control Flow}

    The above properties are somewhat orthogonal, but together they capture a
    standard notion of control flow for a stack: functions keep their code
    separate outside of calls and returns, calls respect explicit entry points,
    and returns respect implicit entry points corresponding to their call.
    A system that enjoys them with regard to a code, a call, a return, and an
    entry map is said to enjoy {\em standard stack control flow} with regard to
    those maps.
